---
title: "EDA"
output: html_document
date: '2022-04-29'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## EDA 
```{r}
library(tidyverse)
library(ggplot2)

library(reshape2)
library(moments)
```

```{r}
df <- read_csv("cleaned_data.csv") %>% 
  mutate_if(~is.numeric(.) && n_distinct(.) <8, as_factor) %>% 
  mutate_if(~is_character(.), as_factor)
```

```{r}
summary(df)
```

```{r}
colSums(is.na(df))
```

```{r}
sum(duplicated(df))
```
remove duplicates later.


### Visualization
```{r}
quants = c("Age", "App_day", "Sche_day", "Sche_hour", "Time_gap")
cate = c("Gender", "Neighbourhood", "Scholarship", "Hipertension", "Diabetes", "Alcoholism", "Handcap", "SMS_received", "App_month", "Sche_year", "Sche_month", "is_workday")
```

```{r}
ggplot(df, aes(x=Age, fill=No_show)) + geom_histogram(position="dodge")
ggplot(df, aes(x=App_day, fill=No_show)) + geom_histogram(position="dodge")
ggplot(df, aes(x=Sche_day, fill=No_show)) + geom_histogram(position="dodge")
ggplot(df, aes(x=Sche_hour, fill=No_show)) + geom_histogram(position="dodge")
ggplot(df, aes(x=Time_gap, fill=No_show)) + geom_histogram(position="dodge")
```

```{r}

ggplot(df, aes(x=Gender)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=Neighbourhood)) + geom_bar(aes(fill=No_show))+ scale_x_discrete(labels = NULL, breaks = NULL) + labs(x = "")
ggplot(df, aes(x=Scholarship)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=Hipertension)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=Diabetes)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=Alcoholism)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=Handcap)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=SMS_received)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=App_month)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=Sche_year)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=Sche_month)) + geom_bar(aes(fill=No_show))
ggplot(df, aes(x=is_workday)) + geom_bar(aes(fill=No_show))



```
Neighbourhood is hard to interpret on histogram due to the large number of neighborhoods. Display each neighborhood's no-show rate (using original dataset to acquire true rate:
```{r}
unbalanced_df <- read_csv("KaggleV2-May-2016.csv")
names(unbalanced_df)[ncol(unbalanced_df)] <- "No_show"
```

```{r}
unbalanced_df %>% 
  group_by(Neighbourhood, No_show) %>%
  summarise(n = n()) %>% 
  mutate(No_show_rate = n /sum(n)) %>% 
  filter(No_show == "Yes") %>% 
  arrange(desc(No_show_rate))
```


```{r}
dim(df)

for (i in quants){
    plot <- ggplot(df, aes_string(df$No_show,i)) +
    geom_violin(aes(fill=No_show))+
    xlab("No_show")
    print(plot)
}
```


```{r}
library(GGally)
```

```{r}
?ggpairs
ggpairs(df[quants],cardinality_threshold=100)
```

### Correlation heatmap 
drop neighbourhood and converting all categorical to numerical:
```{r}
heatmap_df <- df %>% 
  select(-Neighbourhood) %>% 
  mutate_if(~is.factor(.), as.double)

cormat <- round(cor(heatmap_df),2)
melt_cormat <- melt(cormat)

ggplot(melt_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()+
  geom_text(aes(Var2, Var1, label = value), color = "white", size = 3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
```

```{r}
heatmap_df <- df %>% 
  select_if(is.numeric)

heatmap_df <- cbind(heatmap_df,df$No_show)
heatmap_df <- heatmap_df %>% 
  mutate_if(~is.factor(.), as.numeric)

cormat <- round(cor(heatmap_df),2)
melt_cormat <- melt(cormat)

ggplot(melt_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()+
  geom_text(aes(Var2, Var1, label = value), color = "white", size = 3) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
```


# Data preprocessing
### Remove duplicates
```{r}
df <- df[!duplicated(df),]

dim(df)
```
### Skewness correction
```{r}
df %>% 
  select_if(is.numeric) %>% 
  skewness()
```

