---
title: "Preparation"
output: html_document
date: '2022-04-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
```
# Preliminary exploration

### Import dataset
```{r}
df <- read_csv('KaggleV2-May-2016.csv')
```


```{r}
head(df)
### Categorize gender, scholarship, hipertension, diabetes, alcoholism, handcap, SMS_received, No-show
### examine date columns for further processing
### examine Neighbourhood for categorization
### drop IDs
```
```{r}
summary(df)
```

### Feature Engineering
```{r}
df <- read_csv('KaggleV2-May-2016.csv') %>% 
  mutate_if(~is.character(.) && n_distinct(.)<10, as_factor) %>% 
  mutate_if(~is.double(.) && n_distinct(.)<10, as_factor)

head(df)
```
remove ID columns, change column name No-show
```{r}
df <- df %>% 
  select(-c(PatientId, AppointmentID))

names(df)[ncol(df)] <- "No_show"
```


```{r}
summary(df)
```



### Investigate Neighbourhood column
```{r}
df %>% 
  group_by(Neighbourhood) %>% 
  tally(sort=TRUE)
```
Possible methods:
1. categorize, move minority neighbourhoods into "others"
2. Regroup neighbourhoods into larger geographical groups (district, city, etc)

```{r}
df <- df %>% 
  mutate(Neighbourhood = as_factor(Neighbourhood))
```


### Investigate Date columns
```{r}
range(df$AppointmentDay)
range(df$ScheduledDay)
```
year seems unimportant
possible methods:
1. extract month, day, hours for further processing (workday vs holiday, morning/afternoon/night)
2. days between scheduled day and appointment day.

extract month, day from appointment. year, month, day, hour from scheduled.
```{r}
df <- df %>% 
  mutate(App_month = month(AppointmentDay),
         App_day = day(AppointmentDay),
         Sche_year = year(ScheduledDay),
         Sche_month = month(ScheduledDay),
         Sche_day = day(ScheduledDay),
         Sche_hour = hour(ScheduledDay))

head(df)
```

time gap between scheduled day and appointment day
```{r}
df <- df %>% 
  mutate(Time_gap = difftime(AppointmentDay, as.Date(ScheduledDay), units="days")) %>% 
  mutate(Time_gap = as.integer(Time_gap))

sample(df$Time_gap,5)
```
is_workday
```{r}
### All Brazilian holidays between 04-29 to 06-08
holidays = c(ymd("2016-05-01"), ymd("2016-05-08"))

is_holiday = df$AppointmentDay %in% holidays
sum(is_holiday)
# No appointment made in holiday
```

```{r}
is_weekday = wday(df$AppointmentDay)<6
sum(is_weekday)/length(is_weekday)

df <- df %>% 
  mutate(is_workday = as_factor(as.integer(is_weekday)))

levels(df$is_workday)
```


drop date column
```{r}
df <- df %>% 
  select(-c(AppointmentDay, ScheduledDay))

head(df)
```


### Undersampling
```{r}
set.seed(1)

No_show_df <- df[df$No_show=="No",]
No_show_df <- No_show_df[sample(nrow(No_show_df), 22319), ]
Yes_show_df <- df[df$No_show=="Yes",]

cleaned_df <- rbind(No_show_df, Yes_show_df)
```

```{r}
summary(cleaned_df)
```



wrtie a .csv file
```{r}
write_csv(cleaned_df, "cleaned_data.csv")
```

